### 2016-03-07 papers/CAFe/HIPS-paper/coarray-extensions.tex
Added subsection on alternative syntax.

\begin{comment}
It is noted that these extensions to Fortran have been described in terms
explicit language syntax modifying the base language.  CAFe could equivalently have
been described in terms of compiler directives.  For example, subimage task execution could be
written as
\begin{verbatim}
 call relax(U[device]) !$OFP [[device]]
\end{verbatim}
which would allow a standard CAF compiler to compile CAFe programs.  This gives
the flavor of a domain specific language.
\end{comment}

>>>>

\subsubsection{Alternative Syntax}

It is noted that these extensions to Fortran have been described in terms of
explicit language syntax modifying the base language.  CAFe could equivalently have
been described in terms of compiler directives.  For example, subimage task execution could be
written as
\begin{verbatim}
   call relax(U[dev1]) !$OFP [[dev1]]
\end{verbatim}
which would allow a standard CAF compiler to compile CAFe programs.  This syntax usage
would make the CAFe constructs described here an embedded Domain Specific Language.  In a
future version of the implementation of CAFe, we intend to also support compiler directives
in this form.

Alternative syntax for remote execution may be preferable to the double square bracket notation.
While the syntax \texttt{[[ ]]} makes it clear that the programmer intends \emph{remote} execution
to take place, perhaps syntax closer to standard Fortran is better, especially if
standardization of CAFe were to occur.  For example, memory placement could be requested by a
slight modification of the allocation statement
\begin{verbatim}
   allocate(U(M,N)[*], DEVICE=dev1)
\end{verbatim}
and similarly remote execution via a subroutine call could be replaced by
\begin{verbatim}
   call relax(U[dev1]), DEVICE=dev1
\end{verbatim}
However, while the alternative syntax for memory allocation appears very natural, using 
the syntax \texttt{,DEVICE=dev1} to denote remote execution is problematic when applied to
Fortran function calls.  For example, the double square bracket notation in
\begin{verbatim}
   y = sqrt( sin(x)[[dev1]] * 2 )
\end{verbatim}
cannot be replaced with \texttt{,DEVICE=dev1}.
